---
title: "ICA2 Interactive RDBMS data with Quarto"
author: "Tish Lloyd-Lucas D00263110"
format: html
editor: visual
---

---
title: "ICA2 Interactive RDBMS data with Quarto"
author: "Tish Lloyd-Lucas"
format: html
---

## Introduction

This notebook demonstrates three SQL concepts: SELECT with LIKE and OR, SELECT with DISTINCT and ORDER BY, and Subquery with SELECT.

## Setup

```{r}
library(DBI) 
library(RSQLite) 
library(kableExtra) 

# Connect to the SQLite database 
con <- dbConnect(RSQLite::SQLite(), "/Volumes/ANIMATION/DkIT_MSc/YEAR_1__SEMESTER_1/DATA_ANALYTICS_Dr_Muhammad_Adil_Raja/I-CA2_LinearRegression/ICA2_RDBMS_Linear_Regression_TishLL/RDBMS_Linear_Regression/2023_DAIE_ICA_2_TishLL/ICA_2023.sqlite")
```

## 1. SELECT with LIKE and OR

```{r}
# Execute the first query
query1 <- dbGetQuery(con, "SELECT CustomerName, CustomerCity FROM Customers WHERE CustomerName LIKE 'N%' OR CustomerCountry LIKE '%UK';")

# Display results
kable(query1, caption = "Customers whose names start with 'N' or countries contain 'UK'") %>% 
  kable_styling()

```

## 2. SELECT with DISTINCT and ORDER BY

```{r}
# Execute the second query with alphabetical ordering
query2 <- dbGetQuery(con, "SELECT DISTINCT ProjectName FROM Projects ORDER BY ProjectName ASC;")

# Display results
kable(query2, caption = "Project names, ordered alphabetically") %>% 
  kable_styling()

```

## 3. Subquery with SELECT

```{r}
# Execute the query with subquery
query3 <- dbGetQuery(con, "SELECT ProjectName, Budget FROM Projects WHERE Budget > (SELECT AVG(Budget) FROM Projects);")

# Display results
kable(query3, caption = "Projects with above average budget") %>% 
  kable_styling()

```

## Conclusion

This notebook has demonstrated three SQL concepts with practical examples.

## Disconnect from the database

```{r}
dbDisconnect(con)

```
